Discovering Focused Microgenre Communities: Toy example

### 2023-07-31 01_link-distance-matrix.qmd
Updated linkdist function

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    require("dendextend")
    require("fastcluster")
    require("here")   
    require("haven")
    require("tidyverse")
    source(here("Functions", "link.dist.R"))
```

@@

```{r Importing demographic data}
        ed.labs <- c("< High School", "High School/GED", "Associate", "Some College", "Bachelor's", "Postgraduate")
        demog.dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta") %>% 
          dplyr::select(c("id", "age", "agemid", "female", "raceeth", 
                          "nodipdeg", "hsged", "somcol", "aadeg", "bach", "ma", 
                          "docprof", "incmid", "income", "percclass",
                          "region", "occ", "parented", "empstat", "favgen")) %>% 
          as_factor() %>% 
          mutate(educ = if_else(nodipdeg == "yes", 1, 
                                if_else(hsged == "yes", 2, 
                                        if_else(aadeg == "yes", 3, 
                                                if_else(somcol == "yes", 4,
                                                        if_else(bach == "yes", 5,
                                                                if_else(ma == "yes" | docprof == "yes", 6, 1)))))),
                 educ = factor(educ, labels = ed.labs),
                 percclass = fct_recode(percclass,
                                        "Lower Class I.D." = "lower class",
                                        "Workng Class I.D." = "working class",
                                        "Middle/Upper Class I.D." = "middle class",
                                        "Middle/Upper Class I.D." = "upper class"),
                 raceeth = fct_recode(raceeth,
                                      "Asian" = "Asian",
                                      "Black" = "Black",
                                      "Hisp." = "Hispanic",
                                      "White" = "White",
                                      "Mult./Other" = ">1 race",
                                      "Mult./Other" = "other"),
                 female = fct_recode(female, 
                                     "Men" = "male", 
                                             "Women" = "female"),
                 region = fct_recode(region,
                                     "New England" = "New Enlgnad",
                                     "South" = "East South",
                                     "South" = "South Atlantic",
                                     "Southwest" = "West South",
                                     "Midwest" = "Eastern Midwest",
                                     "Midwest" = "Western Midwest",
                                     "Mountain" = "Mountain West",
                                     "West Coast" = "Pacific West"),
                 parented = fct_recode(parented, 
                                       "Parent College? (No)" = "parent no college", 
                                       "Parent College? (Yes)" = "parent college"),
                 empstat_n = factor(as.numeric(empstat)),
                 objclass = case_when(incmid == "5000" | incmid == "15000" ~ 1,
                                      incmid == "25000" | incmid == "35000" ~ 2, 
                                      incmid == "45000" | incmid == "55000" | incmid == "65000" ~ 4,
                                      incmid == "75000" | incmid == "85000" | incmid == "95000" ~ 5,
                                      incmid == "105000" | incmid == "115000" | 
                                        incmid == "125000" | incmid == "135000" |
                                        incmid == "145000" | incmid == "150000" |
                                        incmid == "165000" | incmid == "175000" |
                                        incmid == "185000" | incmid == "195000" |
                                        incmid == "212500" | incmid == "237500" |
                                        incmid == "250,000 or more" ~ 6),
                 objclass = factor(objclass, labels = c("Poor", "Working Class", "Middle Class", "Upper Middle Class", "High Income")),
                 educ_n = as.numeric(educ),
                 college = if_else(educ_n > 4, 1, 0),
                 college = factor(college, labels = c("No Degree", "College")),
                 age_n = as.numeric(age),
                 objclass_n = as.numeric(objclass)
          ) %>% 
          dplyr::select(-c("nodipdeg", "hsged", "somcol", "aadeg", "bach", "ma", "docprof")) %>% 
            mutate(id = as.integer(id)) %>% 
            na.omit() 
```

# Link Clustering
```{r Performing link clustering and saving results}
    ldist.res <- link.dist(x = taste.mat, id = taste.dat$id) #computing link by link distance matrix
    lclust.res <- hclust(ldist.res$dist, method = "ward.D2") #link clustering of link by link distance matrix
    lclust.dend <- as.dendrogram(lclust.res) #transforming link clustering into dendrogram
```

```{r Creating objects for analysis}
    dend.macro <- cut(lclust.dend, h = 8) #cutting at macrogenre level
    a <- data.frame(g.size = sapply(dend.macro$lower, nleaves)) #macrogenres as ordered in dendrogram results
    b <- data.frame(g.size = colSums(taste.mat), g.names = names(colSums(taste.mat))) #original macrogenre sizes
    c <- right_join(a, b) #joining data frames by macrogenre sizes to get macrogenre labels ordered by dendrogram results
    names(dend.macro$lower) <- c$g.names #naming macrogenre dendrogram list
    n.micro <- sapply(dend.macro$lower, function(x) {length(table(cutree(x, h = 3)))}) #number of microgenres per macrogenre
    v <- sapply(sapply(n.micro, function(x) {1:x}), as.numeric)
    micro.g.names <- paste(c$g.names[1], v[[1]], sep = "_")
    for (i in 2:20) {
      micro.g.names <- c(micro.g.names, paste(c$g.names[i], v[[i]], sep = "_"))
        }
    bin.char <- function(x) { 
      if_else(is.na(x), 0, 1) #function to binarize data
        }
    dend.micro <- cut(lclust.dend, h = 3) #cutting at microgenre level
    edge.labs <- data.frame(id = ldist.res$edge.list[, 1],
                            macro.g = ldist.res$edge.list[, 2], 
                            edge.num = 1:nrow(ldist.res$edge.list)) #grabbing people and genre id labels
    micro.g.clust <- lapply(dend.micro$lower, labels) %>%
        lapply(data.frame) %>% 
        bind_rows() #transforming clustering cut results into data frame
    names(micro.g.clust) <- "edge.num" #adding column label
    micro.g.long <- data.frame(micro.g = rep(1:102, sapply(dend.micro$lower, nleaves)))
    micro.g.dat <- bind_cols(micro.g.clust, micro.g.long)
    micro.g.dat <- right_join(micro.g.dat, edge.labs) %>%
      mutate(micro.g.f = factor(micro.g, labels = micro.g.names),
             macro.g.f = factor(macro.g, labels = names(colSums(taste.mat))),
             micro.g.c = as.character(micro.g.f),
             macro.g.c = as.character(macro.g.f)) %>% 
      pivot_wider(id_cols = "id",
                  names_from = "micro.g.c",
                  values_from = "micro.g") %>% 
      mutate(across(is.integer, bin.char)) %>% 
      mutate(id = as.integer(id)) %>% 
      arrange(id)
```

# Saving Objects
```{r Remove garbage and keep important stuff}
    b <- ls()[! ls() %in% c("ldist.res", "lclust.res", "dend.macro", "dend.micro", "lclust.dend", "micro.g.names", "n.micro", "taste.dat", "demog.dat", "micro.g.dat")]
    rm(list = b)
    ls()
```

```{r Saving image}
    save.image(file= here("dat", "lclust-res.RData"))
```

>>>>

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    library(conflicted)
    library(dendextend)
    library(fastcluster)
    library(here)   
    library(haven)
    library(tidyverse)
```

@@

# Link Clustering
```{r Creating link by link distance matrix and saving results}
    source(here("Functions", "link.dist.R"))
    ldist.res <- link.dist(x = taste.mat, id = taste.dat$id)
    b <- ls()[! ls() %in% "ldist.res"]
    rm(list = b)
    ls()
    save.image(file= here("dat", "lclust-res.RData"))
```

### 2023-07-31 02_toy-ex.qmd
Updated linkdist function

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    library("fastcluster")
    library("factoextra")
    library("here")
    library("ggplot2")
    library("sjPlot")
    source(here("R", "Functions", "link.dist.R"))
```

@@

```{r Plotting toy example dendrogram}
    hc.res$labels <- c("P1-G2", "P1-G3", "P2-G4", "P3-G1", "P3-G2", "P3-G3", "P3-G4", "P4-G1", "P4-G4", "P5-G3", "P5-G4")
    p <- fviz_dend(hc.res, k = 4, k_colors = "uchicago", show_labels = TRUE, cex = 1.5, lwd = 2) 
    p <- p + geom_hline(yintercept = 0.75, linetype = "dashed", color = "red", size = 1.2)
    p <- p + labs(y = "", title = "") 
    p <- p + theme(axis.line.y = element_blank(), axis.text.y = element_blank())
    p
    save_plot(here("Plots", "Dend", "link-clust-toy5.png"), 
              width = 40, height = 25)
```

>>>>

# Setup
```{r setup}
    knitr::opts_chunk$set(include=FALSE, echo = TRUE, warning = FALSE, message = FALSE)
    library("conflicted")
    library("fastcluster")
    library("factoextra")
    library("here")
    library("ggplot2")
    library("sjPlot")
    source(here("Functions", "link.dist.R"))
```

@@

```{r Plotting toy example dendrogram}
    hc.res$labels <- c("P1-G2", "P1-G3", "P2-G4", "P3-G1", "P3-G2", "P3-G3", "P3-G4", "P4-G1", "P4-G4", "P5-G3", "P5-G4")
    p <- fviz_dend(hc.res, k = 4, k_colors = "uchicago", show_labels = TRUE, cex = 1.5, lwd = 2) 
    p <- p + geom_hline(yintercept = 0.75, linetype = "dashed", color = "red", linewdith = 1.2)
    p <- p + labs(y = "", title = "") 
    p <- p + theme(axis.line.y = element_blank(), axis.text.y = element_blank())
    p
    save_plot(here("Plots", "Dend", "link-clust-toy5.png"), 
              width = 40, height = 25)
```
